---
title: Methods
author: Author
chapter: 3
in_context: False
knit: iheiddown::chapter_pdf
output: iheiddown::chapter_pdf
---

```{r intialize, echo=FALSE}
# Setting chunk opts

# Libraries
library(skimr)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(tidygraph)
library(ggraph)

# Set paths
project_path <- substr(getwd(), 1, 56)
graph_path <- paste0(project_path, "/graphs/")
ecolex_path <- paste0(project_path, "/data/ECOLEX/Membership/ECOLEX_Membership.rds")
```

\setcounter{chapter}{2}
# Methodology and Data {#methodsdata}

This section will outline the computation of a novel set of centrality measures
which will be used in the subsequent analysis to proxy the relative importance
of countries within the international environmental governance complex and
account for the increasing importance of international policy to tackle
environmental challenges. In what follows, we will describe the construction
of the bipartite membership network of international agreements^[Note that we
interchangeably use bimodal and bipartite as adjectives to qualify the nature
of the original membership network. The original network is both bimodal,
since it has two different types of agreements, and bipartite since it does
not have links between two nodes of the same type.
[@borgatti20092]] and the
extraction of a monopartite cooperation network which defines links between countries
as the number of agreements that have been signed in common. We then
statistically validate this projection by comparing the resulting observed
one cooperation network with a null model. This process is also commonly
known as *backbone extraction*. Finally, we describe the computation of
two centrality indices that will proxy the intensity and the certainty of
international policy. Finally, we will define the empirical panel model that
we will estimate in the next section.


## Defining Hypotheses:

## Constructing international environmental cooperation centrality measures:

We leverage the ECOLEX dataset to construct country level measures of
embeddedness in the international environmental cooperation network
[@unepiucnfaoECOLEXGatewayEnvironmental2022]^[To be more specific, we use a 
pre-scraped version included in the `{manyenviron}` package 
[@hollwayManyenvironEnvironmentalAgreements2021] which was originally used in
@sommerGlobalGovernanceForestry2020]. This dataset lists
environmental agreements between two or more parties and the related
membership actions over the period between 1868 and 2018 and contains over
25000 individual membership actions.^[See Table \@ref(tab:ECOLEX) describing
the structure of the dataset and displaying the first few observations.]
We follow @carattiniWhatDoesNetwork2021
and retain agreements that were signed in the Post-war period starting in
1948. We retain 21270 individual membership actions after filtering out
agreements signed before 1948. After further cleaning the data by excluding
observations on which we do not have identifying information such as the date of
the ratification of the agreement and the date of entry into force or the
country which is the subject of the membership action, we retain 18878
individual membership actions that describe 521 individual international
environmental agreements.

```{r ECOLEX, echo=FALSE, results='asis', fig.pos = 'h'}
load(ecolex_path)
# Remove unsigned
ECOLEX_mems <- ECOLEX_mems %>%
  dplyr::filter(!is.na(Force)) %>%
  dplyr::select(ecolexID, treatyID, CountryID, Title, Beg, End,
                SignatureCountry, Rat, Force, DocType, GeogArea, Subject)
# Create table
kableExtra::kbl(head(ECOLEX_mems, 20),
                caption = "ECOLEX Dataset Head",
                booktabs = T) %>%
  kableExtra::kable_styling(latex_options = c("striped", "scale_down")) %>%
  kableExtra::landscape()
```

### From a Bipartite Membership Network:

The ECOLEX dataset presented above in \@ref(tab:ECOLEX) can be visualized as
an annual series of undirected, unweighted, bipartite/bimodal networks.
This can be done in the form of a yearly incidence matrix $I_{c \times a}^t$ where each row
corresponds to a country and each column to a treaty. $i_{c \times a}$ equals $1$
if country $c$ has ratified agreement $a$ and agreement $a$ has entered into force
and $0$ otherwise. A country can only be a member of a treaty once which implies
that $i_{c \times a} \in \{0,1\}$.

```{=latex}
\begin{equation*}
  \begin{rcases}\displaystyle
    I_{c,a} = \overbrace{\begin{pmatrix}
    i_{1,1}     & i_{1,2}     & i_{1,3}     & \dots     & i_{1,A}           \\
    i_{2,1}     & i_{2,2}     & i_{2,3}     & \dots     & i_{2,A}           \\
    i_{3,1}     & i_{3,2}     & i_{3,3}     & \dots     & i_{3,A}           \\
    \vdots      & \vdots      & \vdots      & \ddots    & \vdots      \\
    i_{C,1}     & i_{C,2}     & i_{C,3}     & \dots     & i_{C,A}
  \end{pmatrix}}^{\text{Agreements}}
  \end{rcases}
    \text{Countries}
\end{equation*}
```

We can also visualize the bipartite/bimodal membership network directly and
distinguish agreements from countries.
The network is depicted in \@ref(fig:ECOLEX_bipartite_network)
represents the membership network with all agreements that entered into force
respectively during the period of 1948 and 1978 and 1948 and 2018. This shows
that the network grew denser as agreements were signed, ratified and entered
into force over time. To explore the intertemporal evolution of
the structure of the network, we will divide this network and construct
yearly snapshots of the membership graph whose incidence matrix we define as
$I_{c \times a}^t$ which is one of the standard methods used in social network
analysis to account for the intertemporal evolution of a network
[@everettDualprojectionApproachTwomode2013].

```{r ECOLEX_bipartite_network, echo=FALSE, fig.pos = 'h'}
ECOLEX_mems_net1 <- ECOLEX_mems %>%
  dplyr::mutate(Year = as.numeric(format(ECOLEX_mems$Force, format = "%Y"))) %>%
  dplyr::filter(Year <= 1978) %>%
  dplyr::select(ecolexID, CountryID) %>%
  migraph::as_tidygraph(twomode = T) %>%
  dplyr::mutate(Type = ifelse(type, "Country", "Agreement"))
ECOLEX_mems_net2 <- ECOLEX_mems %>%
  dplyr::mutate(Year = as.numeric(format(ECOLEX_mems$Force, format = "%Y"))) %>%
  dplyr::select(ecolexID, CountryID) %>%
  migraph::as_tidygraph(twomode = T) %>%
  dplyr::mutate(Type = ifelse(type, "Country", "Agreement"))
# Graph things
layout_bi1 <- ggraph::create_layout(ECOLEX_mems_net1, layout = "igraph",
                                algorithm = "bipartite")
layout_bi2 <- ggraph::create_layout(ECOLEX_mems_net2, layout = "igraph",
                                algorithm = "bipartite")
plot1 <- ggraph::ggraph(ECOLEX_mems_net1, layout = layout_bi1) +
  ggraph::geom_edge_link(alpha = 0.01) +
  ggraph::geom_node_point(aes(shape = Type, color = Type)) +
  ggplot2::labs(title = "Membership Network 1948-1978") +
  ggplot2::theme_void() +
  theme(legend.position = 'bottom')
plot2 <- ggraph::ggraph(ECOLEX_mems_net2, layout = layout_bi2) +
  ggraph::geom_edge_link(alpha = 0.01) +
  ggraph::geom_node_point(aes(shape = Type, color = Type)) +
  ggplot2::labs(title = "Membership Network 1948-2018") +
  ggplot2::theme_void() +
  theme(legend.position = 'bottom',
        )
# Combine plots
gridExtra::grid.arrange(plot1, plot2, ncol = 2)
```

### To a Monopartite Cooperation Network:

While we could further analyze this
network directly, we will transform it into a more easily interpretable
one mode network by using a projection and retaining only significant edges
through a backbone extraction algorithm. We will thus broadly follow
@carattiniWhatDoesNetwork2021 while ensuring that our results are not contingent
on the backbone extraction algorithm since the theoretical underpinnings of
the latter are still somewhat unclear in empirical social network analysis
[@nealComparingAlternativesFixed2021].
We thus begin by creating a naive projection of the membership network. To perform
this naive projection we multiply the yearly incidence matrix $I_{c \times a}^t$
defined above by its transposed $I_{a \times c}^t$ as follows:

$$
I_{c \times a}^t I_{a \times c}^t = P_{c\times c}^t
$$
Where $P_{c\times c}^t$ is the adjacency matrix of the projected cooperation network in
which $p_{ij}^t$ is the number of co-signed agreements between the two countries
that entered into force in the interval $[1948; t]$. These weights capture
the intensity of the environmental cooperation between two countries much like
@newmanScientificCollaborationNetworks2001 captured collaboration within a
bipartite network of scientific collaboration. Concurrently,
it implies that bilateral treaties, ceteris paribus, carry a greater weight
in the cooperation network than multilateral treaties. This is consistent with
the heterogeneous role both play within the environmental governance network.
As highlighted by @hollwayMultilevelEmbeddednessCase2016, bilateral treaties
are akin to contracts between two parties while multilateral treaties serve
"normative [...] law-making tools". While both types may act as policy stringency
signals, bilateral treaties carry more weight through their specific contractual nature
hence developing a greater policy stringency signal for firms. 

We cannot, however, directly analyze this cooperation network as the resulting
*naive* monopartite network, depicted in the first panel of
\@ref(fig:ECOLEX_monopartite), still suffers from two issues. First, nodes
with larger degrees in the original bipartite membership network
(e.g. agreements with a larger number signatories or
countries having signed a larger number of agreements) will bear stronger
edges in the projected network. In other words,
if *country A* and *country B* are members to the same 5 agreements and have
each signed 5 agreements in total, they are conceptually "closer" than
*country C* and *country D* who also share the membership in 5 agreements but
have each signed a total of 50 agreements
[@borgatti2011analyzing; @latapyBasicNotionsAnalysis2008;
@nealBackboneBipartiteProjections2014;
@saraccoInferringMonopartiteProjections2017].
The second issue relates to the fact that *naive* projections may lead to the
emergence of spurious cliques in the projection that are due to a node with a
single connection to the opposite layer in the original bipartite network [@saraccoInferringMonopartiteProjections2017]. To solve these two common issues
we leverage what is known as a backbone extraction algorithm to retain only the
statistically significant edges from the naive projection.

### Correcting the Monopartite Cooperation Network:

We will use the following three distinct backbone extraction methods to ensure
that our results are not the product of the choice of a particular extraction
method alone. All three algorithms, follow the same process in that they
apply a constraint on the degree of the original bipartite network 

#### Fixed Row Model:

#### Fixed Degree Sequence Model:

#### Stochastic Degree Sequence Model:

### Selecting and Computing Network Centrality Measures:

## Other Data Sources:

## The Model:
