---
####################
## Metadata       ##
####################
course: Thesis Proposal
title: |
  International cooperation and technological change
code: ThesisOutline # Course code for filename.
date: "`r paste(Sys.Date())`" # Inserts today's date
authors: Bernhard Bieri
# This is the YAML (YAML Ain't Markup Language) header that includes 
# metadata used to produce the document. 
# Be careful with spacing in this header!
# Two spaces at the start of the line create indentation,
# which embeds the parameter in a hierarchy.

#####################
## Bibliography    ##
#####################
bibliography: [references.bib, packages.bib]
csl: chicago-author-date.csl
nocite: '@*' # Adds all elements in bib file to bibliography

##########################
## Define Knit function ##
##########################
knit: iheiddown::problemset_pdf

#####################
## Output Options  ##
#####################
output: iheiddown::problemset_pdf
link-citations: true
documentclass: article
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Set initial knitr options
knitr::opts_chunk$set(eval = TRUE, echo = FALSE,
                      fig.align = "center",
                      fig.asp = 0.7,
                      dpi = 300,
                      out.width = "80%",
                      fig.pos = "!H",
                      out.extra = "")
# Load packages
library(tidyverse)
```

# Introduction and Motivation:
Climate change occupies a central place in the contemporary social discourse as
becomes apparent by opening any major daily newspaper. The 16th of
February 2022 was no different, the front page of the Financial Times titled
"Biden’s petrol problem: president eyes gas tax cut as pump prices soar"
[@browerBidenPetrolProblem2022] reporting on potential gas tax cuts in
contradiction with Biden's electoral promises. This example highlights the
fact that climate change is at the center the political debate at the national
and the international level.
The motivation for this thesis originated from the Stern Review which
dedicates two chapters on technological innovation and international cooperation
for technological innovation [@sternEconomicsClimateChange2007] as well as
the work I am undertaking on international treaty networks for Prof. Hollway.
While the Stern review provides readers with a good introduction into the
complementary nature of national and international technological policy with
respect to other measures such as carbon pricing, I am interested in
investigating empirically the impact of the structure of the international 
environmental
governance complex on innovation outcomes such as patent applications and patent
citations.

# Litterature review, research question and methodology:

My tentative research question can be therefore stated as follows:

Does international environmental regulation serve as a credible
commitment tool for governments to set policy expectations of firms when making
eco-innovation decisions?

To the best of my knowledge, there are no similar empirical analyses of the
impact of the structure of the international environmental governance complex
on technological innovation. However, there exist two complementary strands of
literature in political science and innovation economics that could be
bridged by this thesis to answer the question at hand.
The political science part of the literature
uses network analysis techniques to understand the structure of multimodal
networks of international agreements and is interested in how political actors
make choices in such a network (see @knokeMultimodalPoliticalNetworks2021).
An example of this could be the determinants
of the choice of a country to join a multilateral environmental agreement for
instance. While these networks are multimodal in nature (e.g. countries are
not directly connected with each other but are signatories to a given treaty),
I plan on using a one-mode country projection of the network of international
environmental agreements thereby reducing its complexity. I would then
extract centrality measures from this one-mode projection for each country and compute yearly averages.
Naturally, depending on the scope of the analysis, the breath of the network
will need to be adjusted since firms active in the energy sector will not see
fisheries regulation as relevant to their R&D decisions. A similar adjustment
will need to be made when considering which countries to include in the
panel (contingent on data availability).

The second strand of literature I considered relevant is the literature on the
economics of climate change and the economics of innovation, with a particular
focus on eco-innovation. The central issue facing policy-makers trying to
mitigate the effects of emissions is one that is characterized by the dual
environmental and knowledge externality. This dual externality leads to both an
underprovision of environmental quality and an underprovision of knowledge
to counteract the first market failure.
Starting with a macro-perspective, one can find
substantial simulation exercises that highlight the complementary role R&D plays
with marked based approaches such as carbon pricing
[@acemogluEnvironmentDirectedTechnical2012; @bosettiRoleTechnologyDiffusion2009].
Acemoglu et al. argue for instance that a combination between carbon prices and
subsidies to mitigate both externalities respectively is less costly than
isolated policy strategies. Furthermore, when taking a more micro-economic
point of view, one can find extensive evidence on the impact of environmental
policy on incentivizing innovation [@veugelersWhichPolicyInstruments2012].
The main body of this subset of the environmental innovation literature focused
on the relationship between a proxy for environmental policy stringency and the
an innovation proxy (patent applications and citations or R&D expenditures). 
The evidence from these studies suggests a positive link between environmental 
policy stringency and rates/quality of innovation e.g. Porter's hypothesis (see @veugelersWhichPolicyInstruments2012 and @martinez-zarzosoDoesEnvironmentalPolicy2019 for more recent evidence on OECD
countries).
However, none of the readings I was able to cover analyse the effect of the
changing structure of the international environmental policy complex (i.e. the
countries within this complex) on these innovation measures. The main idea here
being to test whether more embedded countries generate a greater innovative
output than less embedded ones due to their increased "international
credibility" making their policy commitment more credible and increasing the
expected stringency of environmental policy.

As hinted in the literature review above, the methodology will take the form
of a panel model in which the main variable of interest will be a node/country 
level measure of
"embeddedness" in the international environmental governance network. The
main dependent variable will be a set of proxies for innovative activity in a
given country such as patent counts or R&D expenditures. For validation,
I will also try to include the Community Innovation Survey data or any
similar but more aggregate data into the regressions to validate the results.

# Data:

This subsection will strive to give a preliminary overview of the data I plan
on using to answer the research question stated above and provide the reader
with a descriptive `{skimr}` snapshot.

## Policy network data:

You can find the descriptive snapshots of the [International Environmental 
Agreements](https://iea.uoregon.edu/) and the [Ecolex](https://www.ecolex.org/) databases for 1980 to illustrate the network data.
Each observation in this dataset links a country to a treaty and gives
additional information about their connection such as the nature of the treaty
(bilateral or multilateral), its ratification status, the date of 
signature/ratification/exit, etc.

The main challenge I foresee with this data is to correctly map the scope
of the network to take into account when constructing the relevant network
measures as my variable of interest. Since I will focus primarily
on the energy sector, I will need to detect whether the main focus area of
a given treaty is indeed the energy sector. This will likely go beyond the metadata presented here and involve some level of text analysis on the actual agreements.

```{r, skimrs, warning=FALSE, message=FALSE}
# IEADB
membership_1980IEADB <- manyenviron::memberships$IEADB_MEM %>%
  dplyr::mutate(year = stringr::str_extract(many_ID, "[:digit:]{4}")) %>%
  dplyr::filter(year == "1980")
skimr::skim(membership_1980IEADB)

# ECOLEX
membership_1980ECOLEX <- manyenviron::memberships$ECOLEX_MEM %>%
  dplyr::mutate(year = stringr::str_extract(many_ID, "[:digit:]{4}")) %>%
  dplyr::filter(year == "1980")
skimr::skim(membership_1980ECOLEX)
```

Here is a quick illustration of how one can conceptualize the international 
governance complex as a bimodal network where countries are members of
agreements.

```{r, warning=FALSE, message=FALSE}
# Subset dataset
membership_1980IEADB <- membership_1980IEADB %>%
  dplyr::select(treaty_ID, CountryID, Title) %>%
  tidyr::drop_na() %>%
  migraph::as_tidygraph()

# Plot with migraph
migraph::autographr(membership_1980IEADB) +
  ggplot2::labs(title = "Membership in Environmental Treaties\nSigned in 1980",
       caption = "Source: manyenviron") +
  ggplot2::theme(plot.title = ggplot2::element_text(family = "sans",
                                  size = 18,
                                  hjust = 0.5))
```

## Innovation data:

Innovation, being an indirectly measurable economic phenomenon, will be
measured with a range of proxies. I plan on using country level patent
indicators from the OECD and R&D expenditure as well as their MSTI database.
Alternatively, data could be sourced from PATSTAT directly if needed.

# Format:

I will most certainly use R as the primary data analysis programming
language to execute my analysis since I am most familiar with it and
since there is a very wide ecosystem of packages for network analysis.
This also implies that I will be writing the thesis in R-Markdown using
the `{iheiddown}` dissertation template which will neatly integrate the
code with the analysis. Furthermore, I will be able to use version control 
(GitHub) to track changes throughout the thesis writing and data analysis 
processes which will improve the transparency of the results and ensure
academic honesty. To that effect, the repository will be made public before
the 15th of June 2022.
Note that since my thesis is still in its early stages, the repository
is private but I can grant you or my potential second reader early access
if you would like to.

# Schedule:

The following Gantt chart outlines the projected timeline of the thesis by
breaking down key tasks into smaller goals. This will allow me to track progress over the course of the coming months and will be updated once a month to reflect eventual changes.

```{r, GanttChart, warning=FALSE, message=FALSE}
# Create events data
events <- tibble(wp = c("Litterature Review",
                        "Litterature Review",
                        "Modelling",
                        "Data Analysis",
                        "Data Analysis",
                        "Writing",
                        "Writing",
                        "Deadlines",
                        "Deadlines",
                        "Deadlines"),
                 activity = c("1.1.Thesis Outline",
                              "1.2.Writing Literature Review",
                              "2.1.Creating Theoretical Model(s)",
                              "3.1.Data Cleaning",
                              "3.1.Data Analysis",
                              "4.1.Writing Analysis",
                              "4.2.Integrating Feedback",
                              "5.1 Thesis Outline",
                              "5.2 Intermediate Presentation",
                              "5.3 Final Deadline"),
                 start_date = c(1,
                                1,
                                3,
                                3,
                                3,
                                3,
                                5,
                                2,
                                4,
                                6),
                 end_date = c(2,
                              3,
                              4,
                              4,
                              5,
                              5,
                              6,
                              2,
                              4,
                              6))

# Create gantt chart
ganttrify::ganttrify(project = events,
          project_start_date = "2022-01",
          mark_quarters = FALSE) +
  ggplot2::labs(title = "Master's thesis timeline")
```

# Conclusion and feedback:

Thank you in advance for taking the time to read this thesis outline. I 
would be extremely grateful for any feedback regarding the overall idea, 
tentative methodology, and the format of the thesis.

# References:
<!-- The bibliography is automatically generated here. -->

The references below include the majority of the relevant literature I
read to write this outline.
